---
title: "TestrunMarkdownAfrica"
author: "Sonja Maas"
date: "2023-10-31"
output: pdf_document
---

```{r setup, include=FALSE}
library("terra")
library("rnaturalearth")
library("sf")
library("geodata")
library("rnaturalearth")
library("rnaturalearthdata")
```

# Make a temperature curve and plot for Africa
&nbsp;

### Get the country borders. Other options can be found in the manual.

```{r}
afr <- ne_countries(continent="Africa", scale= "medium", returnclass="sf")
```
&nbsp;

### Plot African boundaries.

```{r, fig.dim=c(6.5,2)}
plot(afr)
```
&nbsp;

### Get the maximum temperatures data.

```{r}
clim <- geodata::worldclim_global(var='tmax', res=10, download=T, path='.')
```
&nbsp;

### Plot the data.

```{r, fig.dim=c(6.5,4)}
plot(clim)
```
&nbsp;

### Check if the project matches (transform "afr" into vector (sf) file) and reproject it.

```{r, fig.dim=c(6.5,6)}
afr.r <- st_transform(afr,st_crs(clim))
clim_afr_crop <- terra::crop(clim, afr.r)
plot(clim_afr_crop)
```
&nbsp;

### Cut the temperature data to the mask of Africa and plot it.

```{r, fig.dim=c(6.5,6)}
clim_afr_mask <- terra::mask(clim_afr_crop, afr.r)
plot(clim_afr_mask)
```
&nbsp;

### Extract the average temperature of Africa and plot it. Other statistics are possible as well.

```{r, fig.dim=c(6.5,6)}
climafr_vect <- terra::extract(clim_afr_mask, afr, mean)
plot(unlist(climafr_vect[,2:13]))
```
